### 2022-09-17 docs/Ideas.rmd
m

Parece que hay una tendencia a buscar clases on line, ademas las universidades estan experimentando bajos enrollments.
Factores socio-enconimicos? mas tiempo para trabajar? correlacion con la inflacion? 
Otras carreras como economia, ingieneria no han tenido una bajada en el numero de solicitudes. Hay un cambio de tendencia en lo que se estudia?.

Preguntas para añadir,

Open questions, such as are you currently working? If so how many hours? Is your job related with your minor/ major?

Do students see language as a useful factor to find a job? 
“ Computer sciences increased by 7.8%, visual and performing arts majors were up 5.7% and psychology majors increased 4%. 
The largest declines among the most popular majors were in liberal arts and sciences, general studies and humanities (-4.2%) and social sciences (-2.3%).”
https://www.forbes.com/sites/michaeltnietzel/2022/05/26/new-report-the-college-enrollment-decline-has-worsened-this-spring/?sh=66eff6ef24e0