### 2022-09-17 docs/Ideas.rmd
m

# Percepciones 
Parece que hay una tendencia a buscar clases on line, ademas las universidades estan experimentando bajos enrollments.
Factores socio-enconimicos? mas tiempo para trabajar? correlacion con la inflacion? 
Otras carreras como economia, ingieneria no han tenido una bajada en el numero de solicitudes. Hay un cambio de tendencia en lo que se estudia?.

# Tipo de preguntas. Number of agreement / describir- open questions

Open questions, such as are you currently working? If so how many hours? Is your job related with your minor/ major?

Do you consider you need to work to cover your university expenses?

Do students see language as a useful factor to find a job? 
“ Computer sciences increased by 7.8%, visual and performing arts majors were up 5.7% and psychology majors increased 4%. 
The largest declines among the most popular majors were in liberal arts and sciences, general studies and humanities (-4.2%) and social sciences (-2.3%).”

https://www.forbes.com/sites/michaeltnietzel/2022/05/26/new-report-the-college-enrollment-decline-has-worsened-this-spring/?sh=66eff6ef24e0

# Background

Have you taken language online class during the pandemic? If so how many?

Have you taken a f2f class after schools started offering them again?

Do you have a preference for online class? Advantages/Disadvantages